Northern Pikeminnow Predation on Juvenile Chinook Salmon in Deadwater Slough, Salmon River, Idaho and Potential Impacts to Local Populations

### 2021-12-22 analysis/paper/paper.Rmd
draft Intro complete

---
title: "Northern Pikeminnow Predation on Juvenile Chinook Salmon and Potential Impacts to Local Populations in the Upper Salmon River, Idaho"
author:
  - Michael W. Ackerman:
      email: mike.ackerman@mthoodenvironmental.com
      institute: [mhe_mccall, biomark]
      correspondence: true
  - Nicholas J. Porter:
      email: Nicholas.Porter@merck.com
      institute: [biomark]
      correspondence: false
  - Kevin E. See:
      email: Kevin.See@dfw.wa.gov
      institute: [biomark, wdfw]
      correspondence: false
  - Michael S. Hall:
      email: Mike.Hall1@merck.com
      institute: [biomark]
      correspondence: false
  - Tara E. Blackman:
      email: tara.blackman@mthoodenvironmental.com
      institute: [mhe_sandy]
      correspondence: false
  - Richard A. Carmichael:
      email: Richard.Carmichael@merck.com
      institute: [biomark]
      correspondence: false
institute:
  - mhe_mccall: Mount Hood Environmental, PO Box 4282, McCall, Idaho, 83638, USA
  - biomark: Biomark, Inc., 705 South 8th Street, Boise, Idaho, 83702, USA
  - wdfw: Washington Department of Fish and Wildife, Fish Program, Science Division, 1111 Washington Street NE, Olympia, Washington, 98501, USA
  - mhe_sandy: Mount Hood Environmental, 39085 Pioneer Boulevard \#100 Mezzanine, Sandy, Oregon, 97055, USA 
date: "`r format(Sys.time(), '%d %B, %Y')`"
always_allow_html: true
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
      number_sections: FALSE
      toc: no
      always_allow_html: true
      fig_width: 6
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
    bookdown::html_document2:
      fig_caption: yes
      number_sections: FALSE
      toc: no
      fig_width: 6
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua      
fontsize: 12pt
mainfont: Times New Roman
bibliography: 
- references.bib      # Kevin's library
- AckermanLibrary.bib # Mike's library
csl: "../templates/american-fisheries-society.csl" # Insert path for the bib-style
abstract: |
  Abstract text here. 
keywords: |
  northern pikeminnow; Chinook salmon; predation; mark-recapture; bioenergetics
highlights: |
  These are the highlights. 
---

@@

The Snake and Columbia rivers drain a network of inland tributaries that provide essnetial spawning and rearing habitat for anadromous Pacific salmon *Oncorhynchus* spp. and steelhead *O. mykiss*. These fishes were historically abundant throughout the drainage basin and consequently have tremendous ecological, cultural, and economic value. Over the last century, salmonid stocks throughout the Snake and Columbia rivers have undergone significant declines related to habitat modification in the freshwater environment (e.g., removal of beavers, river channel simplification, water withdrawals, hydrosystem development) and poor marine conditions, both affecting their survival and recruitment and multiple life stages. Consequently, action agencies have made considerable investments in the rehabilitation of tributary ecosystems and evaulation of factors attributed to the species decline [@Roni2018].

One potentially important, but perhaps under-appreciated source of salmonid mortality is predation on emigrating juveniles by piscivorous fishes. In the lower mainstem Snake and Columbia rivers it is estimated that predation on out-migrating salmonids during peak emigration has a significant negative impact on the overall population and success of population recover [@Fresh2003; @Winther2020]. Dams and reservoirs in the Columbia River are the primary locations associated with high rates of piscine predation on salmonids [@Petersen1994; @Ward1995]. There are generally two mechanisms that explain these high predation zones. First, migration (movement) rates of juvenile salmon are reduced during reservoir passage [@Venditti2000], thereby increasing the time migrating smolts are vulnerable to predation. Second, reservoirs and downstream tailraces associated with dams on the Snake and Columbia rivers create favorable slow-water habitat for predatory fishes known to consume juvenile salmonids, such as the Northern Pikeminnow *Ptychocheilus oregonsis*. Northern Pikeminnow tolerate and thrive in relatively warm, slow water habitats [@Wydoski2003] and consequently have benefited from dams on the Columbia River, becoming abundant predators of salmonid outmigrants [@Knutsen1999]. Indeed, Northern Pikeminnow are estimated to consume 8% (16.4 million) of the approximately 200 million juvenile salmonids emigrating through the lower Snake and Columbia Rivers, annually [@Beamesderfer1996].

Most predation studies have focused on reservoirs [@Murphy2021] and mainstem reaches [@Zimmerman1999; @Tabor1993; @Ward1995; @Shively1996] of the Columbia River basin whereas very little is known about the interaction between upstream, river-dwelling piscivorous fishes and their impacts on salmonid recovery above the reservoir systems [@Rubenson2020]. Upstream habitats containing slower water velocities and other attributes that support piscivorous predators may overlap with essential habitat for some salmonid species and life stages. One such example is Deadwater Slough, an approximately 1.5 kilometer long reach of unnaturally slow and deep water in the Salmon River, Idaho. The Deadwater Slough is located in a river reach containing historically important overwinter rearing and spring emigration habitat for juvenile Chinook Salmon *O. tshawytscha* and is inhabited by piscivorous predators, including Northern Pikeminnow and Smallmouth Bass *Micropterus dolomieu*. Further, the slough is part of the migratory pathway of the endangered Snake River Sockeye Salmon *O. nerka* population [@Axel2015] and several populations of threatened Snake River steelhead.

Quantifying mortality associated with piscine predation in upstream habitats of the Salmon River is of particular interest for spring/summer Chinook Salmon in the Upper Salmon major population group (MPG). The Salmon River was historically the most productive tributary for spring/summer Chinook Salmon in the Columbia River basin [@Nemeth1999], but those populations have become depleted in recent decades. Currently, the Upper salmon MPG supports eight independent, extant spring/summer Chinook Salmon populations, including Salmon River (above Redfish Lake Creek), Valley Creek, Yankee Fork Salmon River, East Fork Salmon River, Salmon River (mainstem below Redfish Lake Creek), Pahsimeroi River, Lemhi River, and North Fork Salmon River (@NOAA2017). The Upper Salmon MPG are stream-type Chinook Salmon and exhibit two distinct migration tactics; downstream rearing (DSR) and natal reach rearing (NRR) [@Copeland2014]. Diversity of migration tactics provides a mechanism for coping with adverse conditions in freshwater rearing and migration environments and buffers against catastrophic events, thereby increasing population resiliency. Because the Deadwater Slough supports rearing for DSR migrants and is part of the migration corridor for NRR migrants, and juveniles from all Upper Salmon populations must rear in or migrate through it, it represents a significant habitat for the Upper Salmon MPG.

Several studies have attempted to examine the downstream movement, distribution, and (apparent) survival of juvenile salmonid emigrants through the Salmon River, including the Deadwater Slough reach. Sockeye Salmon migrating through the mainstem Salmon River during spring were estimated to have substantially reduced survival in the Deadwater Slough reach compared to adjacent reaches [@Axel2015]. Similarly, emigrating DSR Chinook Salmon had a 10% reduction in transition probability through the Deadwater Slough compared to surrounding reaches during the fall and early winter months [@Ackerman2018; @Porter2019]. Low survival in these cases was attributed to low-velocity water delaying movement rates, thereby increasing predation risk.

In this study, we estimated the abundance of Northern Pikeminnow in the Deadwater Slough and its potential impacts to juvenile salmon emigrants, focusing on DSR and NRR Chinook Salmon. We hypothesize that increase densities of piscivorous predators in the Deadwater Slough may explain the reduced survival (or apparent survival) observed for juvenile Chinook Salmon and Sockeye Salmon. To test this, our objectives were three-fold:

@@

<!--
The Upper Salmon River major population group (MPG) supports eight independent, extant spring/summer Chinook Salmon *Oncorhynchus tshawytscha* populations, each providing important ecological, social, and economic benefits to the region. The distinct populations include Salmon River (above Redfish Lake Creek), Valley Creek, Yankee Fork Salmon River, East Fork Salmon River, Salmon River (mainstem below Redfish Lake Creek), Pahsimeroi River, Lemhi River, and North Fork Salmon River [@NOAA2017], all of which have become depleted in recent decades. Reduction in overall population size, and more specifically, declines in survival of juvenile Chinook Salmon have been associated with hydrosystem passage [@Schaller2014], ocean conditions, regional climate [@Welch2021; @Gosselin2021], and downstream predation by native and non-native piscivorous fishes [@Poe1991; @Friesen1999; @Harvey2004]. Anthropogenic changes to freshwater habitat including removal of beavers, river channel simplification, water withdrawals, and the proliferation of non-native species (e.g., non-native coastal rainbow trout *O. mykiss irideus* and brook trout *Salvelinus fontinalis*) have also been attributed to reductions in population size [@IdahoOSCTeam2019]. Moreover, the abundance of returning adults is impacted by ocean and mainstem harvests, poor ocean conditions, and changes to the spawning migration corridor [@NMFS2020BiOp]. Each of these factors have contributed, to varying extents, to reduced adult escapement, the primary metric used to assess population viability. In response to the decline in Upper Salmon Chinook Salmon abundance from the myriad of human activities and associated habitat degradation, action agencies have attempted to improve juvenile survival and adult spawning conditions by investing in rehabilitation of tributary ecosystems as well as investing in the study and evaluation of many factors attributed to the species decline [@Roni2018].

One potentially important, but perhaps under-appreciated source of mortality for Upper Salmon River Chinook Salmon is predation on emigrating juveniles by piscivorous fishes. In the lower mainstem Snake and Columbia rivers it is estimated that predation on out-migrating salmonids during peak emigration has a significant negative impact on the overall population and success of population recovery, most notably in the presence of non-native piscivores [@Fresh2003; @Winther2019]. However, these studies have focused on large rivers and reservoirs of the Snake and Columbia rivers whereas very little is known about the interaction between upstream, river-dwelling piscivorous fishes and their impacts on salmonid recovery above the reservoir systems of the Columbia and Snake river [@Rubenson2020], including the Salmon River. Adding to this, there is little consensus within the literature about the magnitude of predation on salmonids in less altered, free flowing systems, with some suggesting lower rates of predation [@Brown1981; @Buchanan1981] while others suggest more impactful rates of predation [@Tabor1993; @Ward1995; @Shively1996; @Zimmerman1999] that may hinder recovery.

Dams and reservoirs in the Columbia River basin are the primary locations associated with high rates of piscine predation on salmonids [@Petersen1994; @Ward1995]. There are generally two mechanisms that explain these high predation zones. First, movement (swimming) rates of juvenile Chinook Salmon are reduced during reservoir passage [@Venditti2000], thereby increasing the time migrating smolts are vulnerable to predation. Second, reservoirs and downstream tailraces resulting from dams on the Columbia and Snake rivers have created favorable slow-water habitat for several fishes known to consume juvenile Chinook Salmon including both non-native (e.g., Smallmouth Bass *Micropterus dolomieu*, Walleye *Sander vitreus*) and native (e.g., Northern Pikeminnow *Ptychocheilus oregonensis*) species. In Lower Granite Reservoir on the Snake River, juvenile Chinook Salmon are common prey items of Smallmouth Bass during some seasons [@Erhardt2018]. Similarly, non-native Walleye are considered a potentially important source for juvenile Chinook Salmon mortality in the Snake River [@Widener2021], particularly as their abundance has significantly increased surrounding Snake River dam facilities between 2008 and 2020.

Along with non-native species, the Columbia and Salmon rivers support native Northern Pikeminnow, a piscivorous fish tolerating and thriving in relatively warm, slow water habitats. Consequently, they too have benefited from dams on the Columbia River and become abundant predators on salmonid outmigrants [@Knutsen1999]. Indeed, Northern Pikeminnow are estimated to consume 8% (16.4 million) of the approximately 200 million juvenile salmonids emigrating through the lower Snake and Columbia rivers, annually [@Beamesderfer1996]. Predation by piscivorous fishes, including Northern Pikeminnow, may have been a factor in the below-average survival of wild Chinook Salmon in 2020 [@Widener2021].

While significant effort has been devoted to understanding predation-related mortality throughout the hydropower systems of the lower Snake and Columbia rivers, there is a dearth of contemporary data on piscivorous predator populations and little is known about their movement, life history, or impacts upstream of the mainstem habitats [@Gadomski2001]. Consequently, the interaction between upstream, river-dwelling piscivorous fishes and emigrating juvenile salmonids, and its impact on salmonid recovery above reservoir systems of the Columbia and Snake rivers is unknown [@Rubenson2020].

Habitats containing slower water velocities and other attributes that support piscivorous predators are not limited to areas near dams and reservoirs. One such example is the Deadwater Slough, an approximately 1.5 kilometers long reach of the Salmon River, Idaho (Figure \@ref(fig:map)). The slough is located approximately 6 kms downstream from the town of North Fork, Idah with the downstream end located at the confluence with Dump Creek. Around 1987, the failure of a small mining diversion reservoir in the Dump Creek drainage resulted in an erosion event that deposited substantial amounts of sediment at the confluence of the Salmon River, forming an alluvial fan. The influx of sediment created an unnaturally slow and deep section of the river, resulting in "reservoir-like" conditions favorable to predators like Northern Pikeminnow and Smallmouth Bass [@Watkins2015]. Furthermore, the slough lacks hydrological and structureal features that would provide juvenile fishes essential refuge from predation (e.g., large wood or substrate, bank undercuts). As a result, predation of juvenile Chinook Salmon and other local ESA-listed salmonids (Sockeye Salmon *O. nerka*, redband steelhead *O. mykiss gairdneri*) has been cited as a concern for populations in the Upper Salmon River.

The Deadwater Slough is in a reach of the Salmon River believed to be a historically important overwinter rearing and spring emigration area for Chinook Salmon juveniles. Spring/summer Chinook Salmon in the Upper Salmon MPG are stream-type and exhibit two distinct migration tactics; downstream rearing (DSR) and natal reach rearing (NRR) [@Copeland2014a]. The DSR migrants leave the natal spawning area as subyearlings between June and November and typically overwinter in downstream, mainstem habitats until the following spring when they emigrate to the ocean as smolts. Alternatively, NRR migrants remain in their natal spawning areas for approximately one year after emergence until emigration to the ocean as smolts. Diversity of migratory tactics provides a mechanism for coping with adverse conditions in freshwater rearing and migration environments and buffers against catastrophic events, thereby increasing population resiliency. Because the Deadwater Slough both supports rearing for DSR migrants and provides the migration pathway for NRR migrants, it represents a significant habitat for the Upper Salmon MPG.

The slough is also part of the migratory pathway for the endangered Snake River Sockeye Salmon population [@Axel2015] and several populations of threatened Snake River steelhead. Accordingly, the mainstem Salmon River has been the target of recent radio telemetry studies to examine the downstream movement, distribution, and (apparent survival) of juvenile salmonid emigrants. For example, @Axel2015 examined the movement and survival of Sockeye Salmon released in Redfish Lake Creek downstream through the mainstem Salmon River during the spring migration and identified a substantial decrease in estimated survival in the reach spanning the Deadwater Slough. They estimated survival among release groups to be approximately 56% whereas survival through the majority of study reaches was greater than 80%. Further, recent telemetry studies of Chinook Salmon emigration during the fall and early winter months indicated a lower transition probability in the reach spanning the Deadwater Slough, approximately 10% less than surrounding reaches [@Ackerman2018; @Porter2019].

In this study, we estimated the abundance of a piscivorous fish population (Northern Pikeminnow) in the Deadwater Slough and its potential impacts to juvenile salmon emigrants, focusing on DSR and NRR Chinook Salmon. We hypothesize that increased densities of piscivorous predators in the Deadwater Slough may explain the reduced survival (or apparent survival) observed for juvenile Chinook Salmon [@Ackerman2018] and Sockeye Salmon [@Axel2015]. To test this, our objectives for the study were three-fold:

1. Estimate the abundance (or relative abundance) of potential predators in the Deadwater Slough during the peaks of fall (DSR) and spring (NRR) juvenile emigrations.
2. Document predation on juvenile Chinook Salmon during the emigration periods using gastric lavage.
3. Use a bioenergetics approach to estimate consumption potential (grams) of the Northern Pikeminnow population at Deadwater Slough and the percent of consumption consisting of juvenile DSR and NRR Chinook salmon during defined emigration periods.

This research is an essential first step in characterizing conditions through the Deadwater Reach and providing data necessary to discuss potential predation mitigation strategies (if needed), where Northern Pikeminnow removal has proven fruitful in increasing survival through the reservoir systems [@Winther2019]. We follow with a discussion of the various assumptions that went into the mark-recapture and bioenergetics models and how violations of some assumptions may affect overall results and inferences from the study. We also consider how consumption of juvenile Chinook Salmon by predators at Deadwater Slough might impact adult returns, hindering recovery of populations in the Upper Salmon River MPG.
-->

>>>>

---
title: "Northern Pikeminnow Predation on Juvenile Chinook Salmon in Deadwater Slough, Salmon River, Idaho and Potential Impacts to Local Populations"
author:
  - Michael W. Ackerman:
      email: mike.ackerman@mthoodenvironmental.com
      institute: [mhe_mccall, biomark]
      correspondence: true
  - Nicholas J. Porter:
      email: Nicholas.Porter@merck.com
      institute: [biomark]
      correspondence: false
  - Kevin E. See:
      email: Kevin.See@dfw.wa.gov
      institute: [biomark, wdfw]
      correspondence: false
  - Michael S. Hall:
      email: Mike.Hall1@merck.com
      institute: [biomark]
      correspondence: false
  - Tara E. Blackman:
      email: tara.blackman@mthoodenvironmental.com
      institute: [mhe_sandy]
      correspondence: false
  - Richard A. Carmichael:
      email: Richard.Carmichael@merck.com
      institute: [biomark]
      correspondence: false
institute:
  - mhe_mccall: Mount Hood Environmental, PO Box 4282, McCall, Idaho, 83638, USA
  - biomark: Biomark, Inc., 705 South 8th Street, Boise, Idaho, 83702, USA
  - wdfw: Washington Department of Fish and Wildife, Fish Program, Science Division, 1111 Washington Street NE, Olympia, Washington, 98501, USA
  - mhe_sandy: Mount Hood Environmental, 39085 Pioneer Boulevard \#100 Mezzanine, Sandy, Oregon, 97055, USA 
date: "`r format(Sys.time(), '%d %B, %Y')`"
always_allow_html: true
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
      number_sections: FALSE
      toc: no
      always_allow_html: true
      fig_width: 6
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
    bookdown::html_document2:
      fig_caption: yes
      number_sections: FALSE
      toc: no
      fig_width: 6
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua      
fontsize: 12pt
mainfont: Times New Roman
bibliography: 
- references.bib      # Kevin's library
- AckermanLibrary.bib # Mike's library
csl: "../templates/american-fisheries-society.csl" # Insert path for the bib-style
abstract: |
  Abstract text here. 
keywords: |
  Northern Pikeminnow; Chinook Salmon; predation; mark-recapture; bioenergetics
highlights: |
  These are the highlights. 
---

@@

The Snake and Columbia rivers drain a network of inland tributaries that provide essential spawning and rearing habitat for anadromous Pacific salmon *Oncorhynchus* spp. and steelhead *O. mykiss*. These fishes were historically abundant throughout the drainage basin and consequently have tremendous ecological, cultural, and economic value. Over the last century, salmonid stocks throughout the Snake and Columbia rivers have undergone significant declines related to habitat modification in the freshwater environment (e.g., removal of beavers, river channel simplification, water withdrawals, hydrosystem development) and poor marine conditions, both affecting their survival and recruitment at multiple life stages. Consequently, action agencies have made considerable investments in the rehabilitation of tributary ecosystems and evaulation of factors attributed to the species decline [@Roni2018].

One potentially important, but perhaps under-appreciated source of salmonid mortality is predation on emigrating juveniles by piscivorous fishes. In the lower mainstem Snake and Columbia rivers it is estimated that predation on out-migrating salmonids during peak emigration has a significant negative impact on the overall population and success of population recovery [@Fresh2003; @Winther2020]. Dams and reservoirs in the Columbia River are the primary locations associated with high rates of piscine predation on salmonids [@Petersen1994; @Ward1995]. There are generally two mechanisms that explain these high predation zones. First, migration (movement) rates of juvenile salmon are reduced during reservoir passage [@Venditti2000], thereby increasing the time migrating smolts are vulnerable to predation. Second, reservoirs and downstream tailraces associated with dams on the Snake and Columbia rivers create favorable slow-water habitat for predatory fishes known to consume juvenile salmonids, such as the Northern Pikeminnow *Ptychocheilus oregonsis*. Northern Pikeminnow tolerate and thrive in relatively warm, slow water habitats [@Wydoski2003] and consequently have benefited from dams on the Columbia River, becoming abundant predators of salmonid outmigrants [@Knutsen1999]. Indeed, Northern Pikeminnow are estimated to consume 8% (16.4 million) of the approximately 200 million juvenile salmonids emigrating through the lower Snake and Columbia Rivers, annually [@Beamesderfer1996].

Most predation studies have focused on reservoirs [@Murphy2021] and mainstem reaches [@Zimmerman1999; @Tabor1993; @Ward1995; @Shively1996] of the Columbia River basin whereas considerably less is known about the interaction between upstream, river-dwelling piscivorous fishes and their impacts on salmonid recovery above the reservoir systems [@Rubenson2020]. Upstream habitats containing slower water velocities and other attributes that support piscivorous predators may overlap with essential habitat for some salmonid species and life stages. One such example is Deadwater Slough, an approximately 1.5 kilometer long reach of unnaturally slow and deep water in the Salmon River, Idaho. The Deadwater Slough is located in a river reach containing historically important overwinter rearing and spring emigration habitat for juvenile Chinook Salmon *O. tshawytscha* and is inhabited by piscivorous predators, including Northern Pikeminnow and Smallmouth Bass *Micropterus dolomieu*. Further, the slough is part of the migratory pathway of the endangered Snake River Sockeye Salmon *O. nerka* population [@Axel2015] and several populations of threatened Snake River steelhead.

Quantifying mortality associated with piscine predation in upstream habitats of the Salmon River is of particular interest for spring/summer Chinook Salmon in the Upper Salmon major population group (MPG). The Salmon River was historically the most productive tributary for spring/summer Chinook Salmon in the Columbia River basin [@Nemeth1999], but those populations have become depleted in recent decades. Currently, the Upper Salmon MPG supports eight independent, extant spring/summer Chinook Salmon populations, including Salmon River (above Redfish Lake Creek), Valley Creek, Yankee Fork Salmon River, East Fork Salmon River, Salmon River (mainstem below Redfish Lake Creek), Pahsimeroi River, Lemhi River, and North Fork Salmon River [@NOAA2017]. The Upper Salmon MPG are stream-type Chinook Salmon and exhibit two distinct migration tactics; downstream rearing (DSR) and natal reach rearing (NRR) [@Copeland2014]. Diversity of migration tactics provides a mechanism for coping with adverse conditions in freshwater rearing and migration environments and buffers against catastrophic events, thereby increasing population resiliency. Because the Deadwater Slough supports rearing for DSR migrants and is part of the migration corridor for NRR migrants, and juveniles from all Upper Salmon populations must rear in or migrate through it, it represents a significant habitat for the Upper Salmon MPG.

Recent studies have attempted to examine the downstream movement, distribution, and (apparent) survival of juvenile salmonid emigrants through the Salmon River, including the Deadwater Slough reach. Sockeye Salmon migrating through the mainstem Salmon River during spring were estimated to have substantially reduced survival in the Deadwater Slough reach compared to adjacent reaches [@Axel2015]. Similarly, emigrating DSR Chinook Salmon had a 10% reduction in transition probability through the Deadwater Slough compared to surrounding reaches during the fall and early winter months [@Ackerman2018; @Porter2019]. Low survival in these cases was attributed to low-velocity water delaying movement rates, thereby increasing predation risk.

In this study, we estimated the abundance of Northern Pikeminnow in the Deadwater Slough and evaluated potential impacts to juvenile salmon emigrants, focusing on DSR and NRR Chinook Salmon. We hypothesize that high densities of piscivorous predators in the Deadwater Slough may explain the reduced survival (or apparent survival) observed for juvenile Chinook Salmon and Sockeye Salmon. To test this, our objectives were three-fold:

@@

We also consider how consumption of juvenile Chinook Salmo emigrants at Deadwater Slough by Northern Pikeminnow may impact adult returns to the Upper Salmon MPG.
